---
title: "Prepwork - packages"
author: "Lindsay Hayes"
date: "5/6/2019"
output:
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

## Learning Objectives

Install packages necessary for the PG2019 workshop

- Install packages from CRAN
- Install packages from Bioconductor
- Install packages from github

## Assignment

In this assignment, you will install packages one by one in each code block. Then you will check that the packages loaded correctly by running a test line of code.

## TASK 1: Install and Load Packages from CRAN
The most important first step is to run each line individually, not as a code block, so you can check that each package installed correctly. After you installed each package, load the package with the `library()` command to make sure it loads into your environment. 

**Run lines individually:**

- RStudio: click the down arrow next to Run in the panel above. Click "Run Selected Line(s)" 
- MacOS: command + enter
- Windows: ctrl + enter

```{r, message=FALSE, eval=FALSE}
options(repos=c(CRAN="http://cran.us.r-project.org"))
install.packages("ggplot2")
library(ggplot2)
```

```{r, message=FALSE, eval=FALSE}
install.packages("dplyr")
library(dplyr)
```

```{r, message=FALSE, eval=FALSE}
install.packages("gplots")
library(gplots)
```

```{r, message=FALSE, eval=FALSE}
install.packages("devtools")
library(devtools)
```

**TEST:** Check that the packages installed by going to the "Packages" tab in the lower right panel of RStudio and checking that these packages are check marked. 

## TASK 2: Install and Load Packages from Bioconductor
The most important first step is to run each line individually, not as a code block, so you can check that each package installed correctly. After you installed each package, load the package with the `library()` command. Finally, run a test line to make sure it loaded into your environment. Each test should give an output and not an error message. 

Remember to run each line seperately to confirm it installed and loaded correctly. If you run into an error type it into the code block with a "#" so you can come back to it later. 

**When asked if you want to: **

- Update all/some/none? [a/s/n]: n
- Do you want to install from sources the package which needs compilation? (Yes/no/cancel) no

```{r, message=FALSE, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

BiocManager::install("annotatr")     # Install package
library(annotatr)                    # Load package
annotations                          # Test package
```

```{r, message=FALSE, eval=FALSE}
BiocManager::install("rtracklayer")  # Install package
library(rtracklayer)                 # Load package
data("targets")                      # Test package
head(targets)
```

```{r, message=FALSE, eval=FALSE}
BiocManager::install("Homo.sapiens") # Install package
library(Homo.sapiens)                # Load package
Homo.sapiens                         # Test package
```

```{r, message=FALSE, eval=FALSE}
BiocManager::install("bumphunter")   # Install package
library(bumphunter)                  # Load package
dummyData()                          # Test package
```

```{r, message=FALSE, eval=FALSE}
BiocManager::install("RTopper")      # Install package
library(RTopper)                     # Load package
data(exampleData)                    # Test package
dat$dat.affy[1,1]
```

```{r, message=FALSE, eval=FALSE}
biocLite( "reactome.db" )            # Install package
library(reactome.db)                 # Load package
reactome.db                          # Test package
```

```{r, message=FALSE, eval=FALSE}
biocLite( "GenomicRanges" )          # Install package
library(GenomicRanges)               # Load package
GRanges(
    seqnames = Rle(c("chr1", "chr2", "chr1", "chr3"), c(1, 3, 2, 4)),
    ranges = IRanges(101:110, end = 111:120, names = head(letters, 10)),
    strand = Rle(strand(c("-", "+", "*", "+", "-")), c(1, 2, 2, 3, 2)),
    score = 1:10,
    GC = seq(1, 0, length=10))       # Test package
```

## TASK 3: Install and Load Packages from Github
Remember to run each line seperately to confirm it installed and loaded correctly. If you run into an error type it into the code block with a "#" so you can come back to it later. 

```{r, message=FALSE, eval=FALSE}
install_github( "vqv/ggbiplot" )     # Install package
library(ggbiplot)                    # Load package
data("wine")                         # Test package
head(wine)
```